Network Configuration Synthesis with Abstract Topologies

### 2016-11-14 paper/abstraction/paper.tex
wordsmithing 5

The syntax introduced in \S\ref{sec:overview} is just syntactic sugar for a core language based on regular expressions. Figure~\ref{fig:syntax} shows the \sysname syntax.
%
A policy has one or more constraints, each of which is either an aggregation constraint or a path constraint. Path constraints consists of a test on a prefix destination and a list of regular expressions describing preferred network paths. Regular paths are defined over network locations, where each location is either a router inside the operator's network, or an external neighbor connected to the operator's network. Values \In~ and \Out, will match any internal and external location respectively. There are two types of predicates: a test for a concrete prefix $d.d.d.d/[d..d]$ matches a range of IP prefixes (\EG, 10.0.1.0/[24..32]) where metavariable $d$ represents an integer.

@@

A key step for \sysname to generate device configurations is to build a data structure amenable to joint analysis of both the topology and routing policy. Intuitively, \sysname captures both topology and routing constraints by ``intersecting" the finite automata associated with the policy regular expressions with the graph structure of the topology into a new data structure called the Product Graph~\cite{propane} (PG).

The PG compactly encodes all paths that protocol messages should be passed along through the network in order to guarantee that the BGP-speaking devices find policy-compliant paths dynamically.
%
%Although \sysname policies describe the routes that traffic should take through the network, network protocols (including BGP) typically disseminate router information in the opposite direction. For example, a router Y will tell its neighbor X that it has a route to the destination meaning that X can send traffic through Y. For this reason, we start by compiling
%
Because \sysname policies can have a separate routing policy for each predicate matching traffic, we construct one such PG for each predicate.

@@

Paths through the PG represent paths through the topology that protocol messages may use to ensure policy-compliance. For example, in the concrete PG, all messages from N1 are not blocked at spine routers S1 and S2. Paths for a destination learned through N1 will end in an accepting state for the first automaton (\EG, node (T1,1,0)). Similarly, paths for a destination learned through N2 will end in an accepting state for the second automaton.

One interesting observation is that the PG for both the concrete and abstract networks share a similar structure. This leads us to the following observation:

\begin{defn}
\label{def:homomorphism}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete product graph $PG = (G',\mathit{start},P)$ and abstract product graph $PG^A = (G'^A, \mathit{start}^A, P^A)$, then we can construct a new homomorphism $f_{pg} : G' \rightarrow G'^A$ in the following way:
\[ \begin{array}{rcl}
  f_{pg}( \mathit{start} ) & = & \mathit{start}^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{defn}
%
As we will see in \S\ref{sec:generation}, we have proven that our generation strategy commutes with concrete expansion, meaning that we obtain the same results if we expand the abstraction early, or if we defer the expansion until after template generation.
%
Definition~\ref{def:homomorphism} provides some intuition concerning why our strategy for generating templates is correct.
%Abstract product graphs are related to concrete product graphs via a graph homomorphism, just as abstract topologies are related to concrete topologies via a graph homomorphism.

>>>>

The first step for the \sysname compiler is to build the product graph (PG), a data structure that is amenable to joint analysis of the topology and routing policy. As a precursor to this step, we convert the routing policy syntax in \S\ref{sec:overview} to a core language based on regular expressions, shown in Figure~\ref{fig:syntax}.
%
A policy has one or more constraints, each an aggregation or a path constraint. Path constraints have a test on a destination prefix and a list of regular expressions describing network paths. Regular paths are defined over network locations, where each location is either a router inside the operator's network, or an external neighbor connected to that network. Values \In~ and \Out, will match any internal and external location respectively. There are two types of predicates, as shown. A test for a concrete prefix $d.d.d.d/[d..d]$ matches a range of IP prefixes (\EG, 10.0.1.0/[24..32]) where metavariable $d$ represents an integer.

@@

%After converting the routing policy to the regular expression based language, 
We are now ready to generate the PG. Intuitively, the PG captures topology and routing constraints by ``intersecting" the finite automata associated with the policy regular expressions with the graph structure of the topology.
%The PG compactly encodes all paths that protocol messages should be passed along through the network in order to guarantee that the BGP-speaking devices find policy-compliant paths dynamically.
%
%Although \sysname policies describe the routes that traffic should take through the network, network protocols (including BGP) typically disseminate router information in the opposite direction. For example, a router Y will tell its neighbor X that it has a route to the destination meaning that X can send traffic through Y. For this reason, we start by compiling
%
Because there can be a separate routing policy for each predicate matching traffic, we construct one PG for each predicate.

@@

Paths through the PG represent paths through the topology that BGP announcements may use to ensure policy-compliance. For example, in the concrete PG, all messages from N1 are not blocked at spine routers S1 and S2. Paths for a destination learned through N1 will end in an accepting state for the first automaton (\EG, node (T1,1,0)). Similarly, paths for a destination learned through N2 will end in an accepting state for the second automaton.

Interestingly, the concrete and abstract PGs have a similar structure, which leads to the following observation:

\begin{defn}
\label{def:homomorphism}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete product graph $PG = (G',\mathit{start},P)$ and abstract product graph $PG^A = (G'^A, \mathit{start}^A, P^A)$, then we can construct a new homomorphism $f_{pg} : G' \rightarrow G'^A$ in the following way:
\[ \begin{array}{rcl}
  f_{pg}( \mathit{start} ) & = & \mathit{start}^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{defn}
%
As we will see in \S\ref{sec:generation}, we have proven that our generation strategy commutes with concrete expansion, meaning that we obtain the same results if we expand the abstraction early, or if we defer the expansion until after template generation.
%
Definition~\ref{def:homomorphism} provides intuition as to why our strategy for generating templates is correct.
%Abstract product graphs are related to concrete product graphs via a graph homomorphism, just as abstract topologies are related to concrete topologies via a graph homomorphism.
